A Novel Comprehensive Quality Evaluation Model in Automated Semantic Web Service Composition

### 2016-10-20 paper.tex
update

The purpose of web service composition is to accomplish an arbitrarily complex task fulfilling customer's requirement, which could be denoted as composite goal: $G(F(T_{Input}, T_{Output}), NF(T_{QoS}))$. This overall composite goal is demonstrated in two parts. One part is that a given input or input set to get the desired output or output set, it typically refers to a statement in functional requirement ; another part specify the overall acceptable composite quality, which covers aspects in non-functionality. To accomplish the first half goal, two stages are involved in the investigation of this paper : services discovery and service selection. Firstly, service discovery is to find matched web service: $S_{n}(F(S_{Input}, S_{Output}), NF(S_{QoS}))$ from Service Repository: $S \{S_{1}, S_{2},..., S_{n} \}$ with the given Task until the desired output found. It results in more than one solutions without considering optimisation; Service selection is to select web services to reach global best quality($T_{QoS}$). We are aiming to optimise a combined objective function involving both functional and non-functional concerns. Particularly in this paper, two challenges mentioned in Sect. \ref{introduction} are addressed in semantic web composite approach. First, the inputs of each service could be semantically matched by predecessor services in the composition. Second, optimisation in QoS would not be sufficient in many scenarios to make a good decision, if customers are looking for a trade off either in better service quality with counteroffer in QoS or better QoS with less strict functional matchmaking quality.

@@

\subsection{Related Work}
\textbf{Semantic web service matchmaking}. Substantial work \cite{bansal2016generalized,mier2015integrated} on semantic web service composition focused mainly on functional requirements and negect non-functional requirements. This approach utilises Description Logic($DL$) \cite{baader2003description} reasoning between input and output concepts of web services to ensure a semantic matchmaking. Since semantic descriptions are introduced into web service, which is expressed in ontologies, the semantic matchmaking is to considered be a process consisting of seeking similarity of parameters (i.e., input and output) of web services and pairing a mapping between two knowledge representations encoded utilising the ontology \cite{lecue2006formal}. In \cite{shet2012new}, the author surveyed three approaches the similarity measures using taxonomies: one is based on nodes, in which similarity is determined by the information content of the nodes; one is entirely based on edge, where concept distance in a hierarchy structure is evaluated,  and the hybrid approach that combines the previous two methods. A new similarity measure based on edge counting is introduced in a Ontology in \cite{shet2012new}, which extends similarity measure defined by Wu and Palmer \cite{wu1994verbs}. Neighbourhood concepts are considered in their model in Formular \ref{equation1}, where $\lambda$ = 1 (default value as 0).

@@

\section{QoS-aware Semantic Automated Web Service Composition}\label{qswsc_approach}
\begin{figure}[h]
\centerline{
\fbox{
\includegraphics[width=12cm]{overview.pdf}}}
 \caption{Overview of the proposed approach.}
 \label{overview}
\end{figure}
In this paper, we model automated QoS-ware semantic service composition with graphs. PSO has shown its efficiency in solving combinatorial optimisation problems. Therefore, we will propose a PSO based approach, which is considered to be simple and efficient without penalising and repairing comparing to GP \cite{da2014graph}. Fig. \ref{overview} shows the overview of our approach with the several steps included. The composite process is triggered by a composite goal defined in Subsection \ref{problemDes}, which demonstrate customers requirements in terms of functional and non-functional requirement. The previous functional one is defined as $F(T_{Input}, T_{Output})$ , and the later nonfunctional one is the users' accepted level of QoS. Those requirements are initially used to discover all relevant web services and their relationship in semantic matchmaking to get prepared for designing service composition via building graphs. The second step selects relevant web services related to services request. This leads to a shrunken service repository that is used by PSO-based algorithm. In the third step, a service composition graph randomly built up with considering the QoS and matchmaking quality. This graph building is interleaved with semantic matchmaking for services. The output of one service and its associated input of another service is calculated in the matchmaking phase where semantic matchmaking quality are measured which will be explained in Subsection \ref{qulityModel}. The further optimisation of service composition solution is performed over the PSO in Section \ref{pso_algorithm}, which is used to find optimised particle locations mapped back to web service for service-graph building. This approach \cite{da2016particle} is an indirect representation of web service composition that constructing a graph-based composite service composition.

@@

 \caption{Steps of graph-based PSO optimisation technique.}
\label{psoSteps}
\end{algorithm}
The overall algorithm investigated here is a combination of a simple forward decoding PSO \cite{da2016particle} and a graph-building technique \cite{blum1997fast} with our comprehensive quality evaluation model. The idea is to translate the particle location into a service queue as an indirect representation of service composition graph, finding the best fitness of the composite service solution is to discover the optimised location of the particle. In PSO, the dimension of the particle is set up as the same as the number of relevant web services in the shrunken service repository, each of which is mapped to a location in a particle, put services in a queue and sort in ascending order. The second method is to decode a corresponding graph built from that service queue using a forward graph-building technique from a start node to the end node. We select and connect web services to the startNode if this web service can satisfy the given Task input, and an Output Set is initialised and kept updating with outputs from new added web services in the graph, later on, we services are keeping being selected and added if services' inputs could be satisfied by any output in the Output Set, if the Output Set contains the Task's required outputs, we end up graph building. In addition, dangle service nodes and edges should be removed. At last, fitness value can be calculated by the construct-based aggregation.

\subsection{Semantic matchmaking}
The semantic matchmaking is achieved by utilising semantic annotation with OWL and MECE ontologies \cite{bleul2008self} with OWL and XWL format respectively. In OWL, semantics specifies the both concepts and individuals in the forms of classes and instances but limited to taxonomies. The semantic concepts are expressed with the their statements and relationship in RDFs (such as subClassOf ), and the semantic individuals are related to the input and output parameters of web services. The MECE is an extensive annotation to WSDL files, which defines the each parameter of services associated with their semantic individuals. The overall matching making process is demonstrated in Figure \ref{matchmaking}.

@@

The semantic matchmaking transfer a function match between $S_{1}: S_{output} \in C_{1}$ and $S_{2}:S_{output} \in S_{b}$ to a pair of concept match: $match(C_{2}, C_{3})$ defined in Section \ref{semantic Web service Discovery}. The matching mechanism attempts to determine whether correct match criteria between the source concepts of C1 and the target concepts of C2, and the quality of semantic matchmaking for the two its reasoning is also calculated in the quality model in the following Section ref{qulityModel}. The advantage of utilising semantic matchmaking is to enable a reasonable connection between two web services from different service providers to find more service solutions with better QoS.

\subsection{Comprehensive quality model and aggregation matrix}\label{qulityModel}
In this paper, we propose a semantic matchmaking model to evaluate the semantic quality of two matched web services, this fill the gap of current prevailing QoS-aware optimisation consideration in the fully automated composition approaches, which has not considered the quality of matchmaking for service selection.

\textbf{Semantic matchmaking model}. Due to the discretisational characteristics of different match types, which is driven by the cost of  data type integration and manipulation \cite{lecue2009optimizing}, so that match type is considered to be one factor for the semantic quality. Another factor in our proposed model is concept similarity, which could be evaluate based on edge counting method defined by \cite{shet2012new}, a taxonomy based measurement for ontology matching, In this paper, it could be used to estimate the parameter concepts similarity between the output and input of a request for selecting web services with close semantic matching. Therefore, given the quality match type and the similarity of two parameters, the semantic matchmaking quality is defined as Equation \ref{equation4}.$q(s_ {matchType})$ is set up as the same as the quality of the semantic link in literature \cite{lecue2009optimizing} (i.e. either 1 (Exact), 0.75 (Plugin), 0.5 (Subsume) or 0.25 (Intersection)), and $q(s_ {similarity})$ is Equation \ref{equation1}.

@@

\textbf{Comprehensive quality model}. Compared to QoS evaluation model, the comprehensive quality model is established to investigate both functional and non-functional reguirements of composite web service. The non-functional properties typically refer to availability, reliability, cost and response time, which are given by service providers. Consequently, the comprehensive quality model integrated the semantic matchmaking quality and quality of service in Equation \ref{equation5}, which could be broken down into Equation \ref{equation6}.

@@

\textbf{Quality aggregate matrix}.The quality aggregation is defined based on the constructs of composite web services on the criteria of functional and non-functional properties. Typically, two web services are connected by more than one parameter-related concepts. In details, edge-level quality aggregation is determined by the mean of parameter-related concept quality in both semantic match type and similarity respectively, and construct-level quality is further calculated based on the edge-level quality aggregation following the rules in Table \ref{table1}, which is the quality aggregate matrix for semantic web service composition. The calculation of QoS aggregation is similar to \cite{cardoso2004quality} while semantic matchmaking aggregation is adopted from \cite{lecue2009optimizing}.

% Please add the following required packages to your document preamble:
% \usepackage{multirow}
\begin{table}[]
\centering
\caption{Quality aggregate matrix for semanic web service composition}
\label{table1}
\begin{tabular}{|c|c|c|c|l|}
\hline
\multicolumn{3}{|c|}{Composition Construct}                                      & Sequence                             & Parallel \\ \hline
\multirow{5}{*}{QualityFactors} & \multirow{2}{*}{Functional}    & $Q(s_ {mt})$  &$\prod_{n=1}^{m} q(s_ {mt})$          &  $\prod_{n=1}^{m} q(s_ {mt})$ \\ \cline{3-5}
                                &                                & $Q(s_ {s})$  & $(\sum_{n=1}^m q(s_ {s}))/m$        &  $(\sum_{n=1}^m q(s_ {s}))/m$  \\ \cline{2-5}   
                                & \multirow{4}{*}{NonFunctional} & $Q(s_{a})$    & $\prod_{n=1}^{m} q(s_a)$             &  $\prod_{n=1}^{m} q(s_a)$ \\ \cline{3-5} 
                                &                                & $Q(s_{r})$    & $\prod_{n=1}^{m} q(s_r)$             &  $\prod_{n=1}^{m} q(s_r)$ \\ \cline{3-5} 
                                &                                & $Q(s_{c})$    & $\sum_{n=1}^m q(s_ {c})$             &  $\sum_{n=1}^m q(s_ {c})$ \\ \cline{3-5} 
                                &                                & $Q(s_{t})$    & $\sum_{n=1}^m q(s_ {t})$             &  $max(q(s_ {t}))$ \\ \hline
\end{tabular}
\end{table}

@@

We run the experiment under computing grid comprising of almost 170 NetBSD (Unix operating system) workstations operated by the Sun Grid Engine. Experimentation was done using the composite service approach explained in Section \ref{qswsc_approach}, and evaluated in the comprehensive quality model introduced in Section \ref{qulityModel}. The parameters were chosen based on general settings from \cite{shi2001particle}, In PSO approach, 30 particles evolved in 100 generations in the searching space, in which problem dimension's size equals to the relevant service size shown in graph. we run 30 times independently for each dataset. In addition, weights setting are more flexible defined by users' preferences. Simply, we configure weights of fitness function as same leverage for balancing functional side and nonfunctional side. Therefore, W1 and W2 are equals to 0.1 and 0.4 entirety 0.5,  and W3, W4, W5, W6 are all set to 0.125 summed 0.5.

>>>>

The purpose of web service composition is to accomplish an arbitrarily complex task fulfilling customer's requirement, which could be denoted as composite goal: $Comp.G(F(T_{Input}, T_{Output}), NF(T_{QoS}))$. This overall composite goal is demonstrated in two parts. One part is that a given input or input set to get the desired output or output set, it typically refers to a statement in functional requirement ; another part specify the overall acceptable composite quality, which covers aspects in non-functionality. To accomplish the first half goal, two stages are involved in the investigation of this paper : services discovery and service selection. Firstly, service discovery is to find matched web service: $S_{n}(F(S_{Input}, S_{Output}), NF(S_{QoS}))$ from Service Repository: $S \{S_{1}, S_{2},..., S_{n} \}$ with the given Task until the desired output found. It results in more than one solutions without considering optimisation; Service selection is to select web services to reach global best quality($T_{QoS}$). We are aiming to optimise a combined objective function involving both functional and non-functional concerns. Particularly in this paper, two challenges mentioned in Sect. \ref{introduction} are addressed in semantic web composite approach. First, the inputs of each service could be semantically matched by predecessor services in the composition. Second, optimisation in QoS would not be sufficient in many scenarios to make a good decision, if customers are looking for a trade off either in better service quality with counteroffer in QoS or better QoS with less strict functional matchmaking quality.

@@

\subsection{Related Work} \label{relatedWork}
\textbf{Semantic web service matchmaking}. Substantial work \cite{bansal2016generalized,mier2015integrated} on semantic web service composition focused mainly on functional requirements and negect non-functional requirements. This approach utilises Description Logic($DL$) \cite{baader2003description} reasoning between input and output concepts of web services to ensure a semantic matchmaking. Since semantic descriptions are introduced into web service, which is expressed in ontologies, the semantic matchmaking is to considered be a process consisting of seeking similarity of parameters (i.e., input and output) of web services and pairing a mapping between two knowledge representations encoded utilising the ontology \cite{lecue2006formal}. In \cite{shet2012new}, the author surveyed three approaches the similarity measures using taxonomies: one is based on nodes, in which similarity is determined by the information content of the nodes; one is entirely based on edge, where concept distance in a hierarchy structure is evaluated,  and the hybrid approach that combines the previous two methods. A new similarity measure based on edge counting is introduced in a Ontology in \cite{shet2012new}, which extends similarity measure defined by Wu and Palmer \cite{wu1994verbs}. Neighbourhood concepts are considered in their model in Formular \ref{equation1}, where $\lambda$ = 1 (default value as 0).

@@

\section{QoS-aware Semantic Automated Web Service Composition}\label{qswsc_approach}
\begin{figure}[h]
\centerline{
\fbox{
\includegraphics[width=12cm]{overview.pdf}}}
 \caption{Overview of the proposed approach.}
 \label{overview}
\end{figure}
In this paper, we model automated QoS-ware semantic service composition with graphs. PSO has shown its efficiency in solving combinatorial optimisation problems. Therefore, we will propose a PSO based approach, which is considered to be simple and efficient without penalising and repairing comparing to GP \cite{da2014graph}. Fig. \ref{overview} shows the overview of our approach with the five steps included. Step 1:The composite process is triggered by a composite goal defined in Subsection \ref{problemDes}, which describes customers requirements in terms of functional and non-functional requirement. The previous functional one is defined as $F(T_{Input}, T_{Output})$ , and the later nonfunctional one is the users' accepted level of QoS. Step 2: Those requirements are initially used to discover all relevant web services and their relationship in semantic matchmaking to get prepared for designing service composition via building graphs. This step selects relevant web services related to services request. This leads to a shrunken service repository that is used by PSO-based algorithm. Step 3: a service composition graph is randomly built up with considering the QoS and matchmaking quality. This graph building is interleaved with semantic matchmaking for services. The output of one service and its associated input of another service is calculated in the matchmaking phase where semantic matchmaking quality are measured which will be explained in Subsection \ref{qualityModel}. Step 4,5: The optimisation of service composition solution is performed over the PSO in Section \ref{pso_algorithm}, which is used to find optimised particle locations mapped back to web service for service-graph building. This approach \cite{da2016particle} is an indirect representation of web service composition that constructing a graph-based composite service composition.

@@

 \caption{Steps of graph-based PSO optimisation technique.}
\label{psoSteps}
\end{algorithm}
The overall algorithm investigated here is a combination of a simple forward decoding PSO \cite{da2016particle} with our comprehensive quality evaluation model. The idea is to translate the particle location into a service queue as an indirect representation of service composition graph, finding the best fitness of the composite service solution is to discover the optimised location of the particle. In PSO, the dimension of the particle is set up as the same as the number of relevant web services in the shrunken service repository, each of which is mapped to a location in a particle, put services in a queue and sort in ascending order. Later on, we decode a corresponding graph built from that service queue using a forward graph-building technique from a start node to the end node. We select and connect web services to the startNode if this web service can satisfy the given Task input, and an Output Set is initialised and kept updating with outputs from new added web services in the graph, later on, we services are keeping being selected and added if services' inputs could be satisfied by any output in the Output Set, if the Output Set contains the Task's required outputs, we end up graph building. In addition, dangle service nodes and edges should be removed. At last, fitness value can be calculated by the construct-based aggregation.

\subsection{Semantic matchmaking}
The semantic matchmaking is achieved by utilising semantic annotation with OWL and OWL-S. In this paper, we use MECE ontologies \cite{bleul2008self} XWL format, which could be considered to be an alternative form of ontology for services. In OWL, semantics specifies the both concepts and individuals in the forms of classes and instances but limited to taxonomies. The semantic concepts are expressed with the their statements and relationship in RDFs (such as subClassOf ), and the semantic individuals are related to the input and output parameters of web services. The MECE is an extensive annotation to WSDL files, which defines the each parameter of services associated with their semantic individuals. The overall matching making process is demonstrated in Figure \ref{matchmaking}.

@@

The semantic matchmaking transfer a function match between $S_{1}: S_{output} \in C_{1}$ and $S_{2}:S_{output} \in S_{b}$ to a pair of concept match: $match(C_{2}, C_{3})$ defined in Section \ref{semantic Web service Discovery}. The matching mechanism attempts to determine whether correct match criteria between the source concepts of C1 and the target concepts of C2, and the quality of semantic matchmaking for these two concepts is calculated in the quality model in subSection \ref{qualityModel}. The advantage of utilising semantic matchmaking quality is to find a more closely requested service output with acceptable level of QoS.

\subsection{Comprehensive quality model and aggregation matrix}\label{qualityModel}
In this paper, we propose a semantic matchmaking model to evaluate the semantic quality of two matched web services, this fill the gaps in subsection \ref{relatedWork} of current prevailing QoS-aware optimisation, where quality of matchmaking for service selection are not considered.

\textbf{Semantic matchmaking model}. Due to the discretisational characteristics of different match types, which is driven by the cost of  data type integration and manipulation \cite{lecue2009optimizing}. Match type is considered to be one factor for the semantic quality. Another factor in our proposed model is concept similarity, which could be evaluate based on edge counting method defined by \cite{shet2012new}, a taxonomy based measurement for ontology matching, In this paper, it could be used to estimate the parameter concepts similarity between the output and input of a request for selecting web services with close semantic matching. Therefore, given the quality match type and the similarity of two parameters, the semantic matchmaking quality is defined as Formula \ref{equation4}, where $q(s_ {matchType})$ is set up as the same as the quality of the semantic link in literature \cite{lecue2009optimizing} as 1 (Exact), 0.75 (Plugin), 0.5 (Subsume) or 0.25 (Intersection) while $q(s_ {similarity})$ is calculated utilising formula \ref{equation1} in subsection \ref{relatedWork}.

@@

\textbf{Comprehensive quality model}. Compared to QoS evaluation model, the comprehensive quality model is established to investigate both functional and non-functional requirements of composite web services. The non-functional properties typically refer previously discussed QoS, which are given by service providers. Consequently, the comprehensive quality model integrated the semantic matchmaking quality $q(s_{smq})$ in Formula \ref{equation5}, which could be further broken down into Formula \ref{equation6}.

@@

\textbf{Quality aggregate matrix}.The quality aggregation is defined based on the constructs of composite web services on the criteria of functional and non-functional properties. Typically, web services could have more than one concept-related parameters. Regarding semantic matchmaking part, edge-level quality aggregation is determined by the mean of concept-related parameter quality in both semantic match type and similarity respectively, and construct-level quality is further calculated based on the edge-level quality aggregation following the rules in Table \ref{table1}, which is the quality aggregate matrix for semantic web service composition. Semantic matchmaking aggregation considers two different factor aggregations: the matching type quality aggregation is calculated by the product function of involved edge-level matching type value, and the similarity quality average value of all edge-level similarity quality. The nonfunctional (QoS) aggregation calculation is same to \cite{cardoso2004quality}.

% Please add the following required packages to your document preamble:
% \usepackage{multirow}
\begin{table}[]
\centering
\caption{Quality aggregate matrix for semantic web service composition}
\label{table1}
\begin{tabular}{|c|c|c|c|l|}
\hline
\multicolumn{3}{|c|}{Composition Construct}                                      & Sequence                             & Parallel \\ \hline
\multirow{5}{*}{QualityFactors} & \multirow{2}{*}{Functional}    & $Q(s_ {mt})$  &$\prod_{n=1}^{m} q(s_ {mt})$          &  $\prod_{n=1}^{m} q(s_ {mt})$ \\ \cline{3-5}
                                &                                & $Q(s_ {s})$  & $(\sum_{n=1}^m q(s_ {s}))/m$        &  $(\sum_{n=1}^m q(s_ {s}))/m$  \\ \cline{2-5}   
                                & \multirow{4}{*}{NonFunctional} & $Q(s_{a})$    & $\prod_{n=1}^{m} q(s_a)$             &  $\prod_{n=1}^{m} q(s_a)$ \\ \cline{3-5} 
                                &                                & $Q(s_{r})$    & $\prod_{n=1}^{m} q(s_r)$             &  $\prod_{n=1}^{m} q(s_r)$ \\ \cline{3-5} 
                                &                                & $Q(s_{c})$    & $\sum_{n=1}^m q(s_ {c})$             &  $\sum_{n=1}^m q(s_ {c})$ \\ \cline{3-5} 
                                &                                & $Q(s_{t})$    & $\sum_{n=1}^m q(s_ {t})$             &  $max(q(s_ {t}))$ \\ \hline
\end{tabular}
\end{table}

@@

We run the experiment under computing grid comprising of almost 170 NetBSD (Unix operating system) workstations operated by the Sun Grid Engine. Experimentation was done using the composite service approach explained in Section \ref{qswsc_approach}, and evaluated in the comprehensive quality model introduced in Section \ref{qualityModel}. The parameters were chosen based on general settings from \cite{shi2001particle}, In PSO approach, 30 particles evolved in 100 generations in the searching space, in which problem dimension's size equals to the relevant service size shown in graph. we run 30 times independently for each dataset. In addition, weights setting are more flexible defined by users' preferences. Simply, we configure weights of fitness function as same leverage for balancing functional side and nonfunctional side. Therefore, W1 and W2 are equals to 0.1 and 0.4 entirety 0.5,  and W3, W4, W5, W6 are all set to 0.125 summed 0.5.
